Equations & R Code

### 2017-03-13 manuscript/coexistence-stability_SI.Rmd
done with first rewrite

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{S1-\arabic{figure}}  
\renewcommand\thetable{S1-\arabic{table}}  
\renewcommand\thesection{Section SI.\arabic{section}}

@@

In the main text we did explore some parameter space by conducting simulations across gradients of environmental variance (whether the cue or the resource), correlations of species' responses to environmental conditions, and competitive asymmetries.
In many cases, altering model parameters, or making them asymmetric among species, makes coexistence more difficult. 
For example, as we show in Figure 3 in the main text, imposing competitive hierarchies makes it more difficult for species to coexist, but it does not impact our conclusion that a positive diversity--ecosystem variability relationship is possible.
We imposed competitive hierarchies by making the live-to-dormant biomass fractions unique for each species, but any parameter that controls population growth would do the same.
For example, if we make dormant mortality rates assymetric among species we achieve similar results as shown in Figure 3 (Fig. S1-1).

@@

This result (Fig. S1-x and Figure 3 in the main text) reinforces the notion that our conclusion about a positive diversity--ecosystem variability relationship is contingent upon the regional community consisting of species that *can* coexist at some level of observable environmental variance.
In other words, if there is no species that can take advantage of more variable conditions, then fluctuation-dependent coexistence is not possible and neither are the results we present in the main text.

@@

Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability (richness increases along x-axes within panels); (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig. S1-2).
Although the qualitative patterns are similar, increasing the dormant mortality rate has two interesting effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by reducing mean population size a little bit and by reducing the temporal standard deviation a lot, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig. S1-2).
The second result occurs because population fluctuations are reduced when very little biomass can be activated to the live stage at the beginning of each season.
That is, in terms of an annual plant, germination of live biomass is always a fraction of a very small number.
The third result is a consequence of the second: populations are not fluctuating that much, so total *CV* can only be reduced by so much with the addition of another species.

@@

As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. S1-3).
Our general conclusion remains, however, that positive diversity--ecosystem variability relationships are possible.
However, the occurence a of positive diversity--ecosystem variability relationship depends on amount of resource variability being sampled over and the traits of individual species (e.g., resource uptake curves). 
In other words, the patterns we report in the main text require that fluctuation-dependent species coexistence be possible given the environmental conditions and species' traits.

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/regional_diversity_stability_storage_effect_8species.png}
  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect. Results are from simulations where environmental variance determines the number species that coexist in a community (e.g., a ``regional'' relationship. Colored points show results from individual simulations and the gray points with connecting line show the mean at each level of environmental variance.}
\end{figure}

From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure S1-4).
However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure S1-4).
Thus, our results provide novel theoretical explanations for positive and flat diversity--ecosystem variability relationships.

@@

>>>>

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{SI-\arabic{figure}}  
\renewcommand\thetable{SI-\arabic{table}}  
\renewcommand\thesection{Section SI.\arabic{section}}

@@

In the main text we did explore some parameter space by conducting simulations across gradients of environmental variance (whether the cue or the resource), correlations of species' responses to environmental conditions, and competitive asymmetries.
In many cases, altering model parameters, or making them asymmetric among species, makes coexistence more difficult. 
For example, as we show in Figure 3 in the main text, imposing competitive hierarchies makes it more difficult for species to coexist, but it does not impact our conclusion that a positive diversity--ecosystem variability relationship is possible.
We imposed competitive hierarchies by making the live-to-dormant biomass fractions unique for each species, but any parameter that controls population growth would do the same.
For example, if we make dormant mortality rates assymetric among species we achieve similar results as shown in Figure 3 (Fig. SI-1).

@@

This result (Fig. SI-1 and Figure 3 in the main text) reinforces the notion that our conclusion about a positive diversity--ecosystem variability relationship is contingent upon the regional community consisting of species that *can* coexist at some level of observable environmental variance.
In other words, if there is no species that can take advantage of more variable conditions, then fluctuation-dependent coexistence is not possible and neither are the results we present in the main text.

@@

Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability (richness increases along x-axes within panels); (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig. SI-2).
Although the qualitative patterns are similar, increasing the dormant mortality rate has two interesting effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by reducing mean population size a little bit and by reducing the temporal standard deviation a lot, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig. SI-2).
The second result occurs because population fluctuations are reduced when very little biomass can be activated to the live stage at the beginning of each season.
That is, in terms of an annual plant, germination of live biomass is always a fraction of a very small number.
The third result is a consequence of the second: populations are not fluctuating that much, so total *CV* can only be reduced by so much with the addition of another species.

@@

As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. SI-3).
Our general conclusion remains, however, that positive diversity--ecosystem variability relationships are possible.
However, the occurence a of positive diversity--ecosystem variability relationship depends on amount of resource variability being sampled over and the traits of individual species (e.g., resource uptake curves). 
In other words, the patterns we report in the main text require that fluctuation-dependent species coexistence be possible given the environmental conditions and species' traits.

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/SI_storage_effect_eightspp_local_regional.png}
  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect. The left panel shows results from simulations where environmental cue variance determines the number species that coexist in a community. The right panel shows results from simulations where environmental cuie variance is fixed at a level that allows coexistence of all eight species, but species are removed to manipulate diversity. The left-hand panel represents a ``regional'' diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a ``local'' diversity--ecosystem variability relationship. Points are jittered within discrete richness values for visual clarity.}
\end{figure}

From one to four species, the relationship is as presented in the main text: total community *CV* increases approximately linearly with environmental variability because (1) environmental variability promotes species coexistence \emph{and} (2) environmental variability causes populations fluctuations to increase (Figure SI-4).
However, after four species, the relationship saturates --- species additions due to coexistence by the storage effect completely buffer ecosystem variability from further increases in environmental variability (Figure SI-4).
Thus, our results provide novel theoretical explanations for positive and flat diversity--ecosystem variability relationships.

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{./components/SI_temp_means_sds_local.png}
  \caption{Temporal means and standarad deviations of total community biomass at each level of species richness. Values come from the simulations represented in the main text Figure 2B (storage effect) and Figure 2D (relative nonlinearity). These plots show how the components of the coefficient of variability change with species richness at a given level of environmental variability (e.g., a ``local'' diversity--ecosystem variability relationship).}
\end{figure}

### 2017-03-13 manuscript/coexistence-stability_SI.tex
done with first rewrite

This result (Fig. S1-x and Figure 3 in the main text) reinforces the
notion that our conclusion about a positive diversity--ecosystem
variability relationship is contingent upon the regional community
consisting of species that \emph{can} coexist at some level of
observable environmental variance. In other words, if there is no
species that can take advantage of more variable conditions, then
fluctuation-dependent coexistence is not possible and neither are the
results we present in the main text.

@@

Our results are qualitatively similar: (1) increasing environmental
variation allows more species to coexist, creating a positive
relationship between species richness and ecosystem variability; (2) and
at any given level of environmental variability, it is always better to
have more species, creating a within-site negative relationship between
species richness and ecosystem variability (Fig. S1-2). Although the
qualitative patterns are similar, increasing the dormant mortality rate
has two interesting effects: (1) it makes coexistence more difficult,
(2) it reduces the absolute value of ecosystem \emph{CV} by reducing
mean population size a little bit and by reducing the temporal standard
deviation a lot, and (3) it weakens the buffering effect of additional
species (compare the spread of the lines between top and bottom panels
of Fig. S1-2). The second result occurs because population fluctuations
are reduced when very little biomass can be activated to the live stage
at the beginning of each season. That is, in terms of an annual plant,
germination of live biomass is always a fraction of a very small number.
The third result is a consequence of the second: populations are not
fluctuating that much, so total \emph{CV} can only be reduced by so much
with the addition of another species.

@@

As expected, lower mean levels of resource make coexistence more
difficult because resource acquisitive species are never favored (Fig.
S1-3). Our general conclusion remains, however, that positive
diversity--ecosystem variability relationships are possible. However,
the occurence a of positive diversity--ecosystem variability
relationship depends on amount of resource variability being sampled
over and the traits of individual species (e.g., resource uptake
curves). In other words, the patterns we report in the main text require
that species coexistence be possible given the environmental conditions.

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=3in]{./components/regional_diversity_stability_storage_effect_8species.png}
  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect. Results are from simulations where environmental variance determines the number species that coexist in a community (e.g., a ``regional'' relationship. Colored points show results from individual simulations and the gray points with connecting line show the mean at each level of environmental variance.}
\end{figure}

From one to four species, the relationship is as presented in the main
text: total community \emph{CV} increases approximately linearly with
environmental variability because (1) environmental variability promotes
species coexistence \emph{and} (2) environmental variability causes
populations fluctuations to increase (Figure S1-4). However, after four
species, the relationship saturates --- species additions due to
coexistence by the storage effect completely buffer ecosystem
variability from further increases in environmental variability (Figure
S1-4). Thus, our results provide novel theoretical explanations for
positive and flat diversity--ecosystem variability relationships.

>>>>

This result (Fig. SI-1 and Figure 3 in the main text) reinforces the
notion that our conclusion about a positive diversity--ecosystem
variability relationship is contingent upon the regional community
consisting of species that \emph{can} coexist at some level of
observable environmental variance. In other words, if there is no
species that can take advantage of more variable conditions, then
fluctuation-dependent coexistence is not possible and neither are the
results we present in the main text.

@@

Our results are qualitatively similar: (1) increasing environmental
variation allows more species to coexist, creating a positive
relationship between species richness and ecosystem variability
(richness increases along x-axes within panels); (2) and at any given
level of environmental variability, it is always better to have more
species, creating a within-site negative relationship between species
richness and ecosystem variability (Fig. SI-2). Although the qualitative
patterns are similar, increasing the dormant mortality rate has two
interesting effects: (1) it makes coexistence more difficult, (2) it
reduces the absolute value of ecosystem \emph{CV} by reducing mean
population size a little bit and by reducing the temporal standard
deviation a lot, and (3) it weakens the buffering effect of additional
species (compare the spread of the lines between top and bottom panels
of Fig. SI-2). The second result occurs because population fluctuations
are reduced when very little biomass can be activated to the live stage
at the beginning of each season. That is, in terms of an annual plant,
germination of live biomass is always a fraction of a very small number.
The third result is a consequence of the second: populations are not
fluctuating that much, so total \emph{CV} can only be reduced by so much
with the addition of another species.

@@

As expected, lower mean levels of resource make coexistence more
difficult because resource acquisitive species are never favored (Fig.
SI-3). Our general conclusion remains, however, that positive
diversity--ecosystem variability relationships are possible. However,
the occurence a of positive diversity--ecosystem variability
relationship depends on amount of resource variability being sampled
over and the traits of individual species (e.g., resource uptake
curves). In other words, the patterns we report in the main text require
that fluctuation-dependent species coexistence be possible given the
environmental conditions and species' traits.

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/SI_storage_effect_eightspp_local_regional.png}
  \caption{Variability of total community biomass as a function of species richness when coexistence is maintained by the storage effect. The left panel shows results from simulations where environmental cue variance determines the number species that coexist in a community. The right panel shows results from simulations where environmental cuie variance is fixed at a level that allows coexistence of all eight species, but species are removed to manipulate diversity. The left-hand panel represents a ``regional'' diversity--ecosystem variability relationship across a natural diversity gradient, whereas the right-hand panel represents a ``local'' diversity--ecosystem variability relationship. Points are jittered within discrete richness values for visual clarity.}
\end{figure}

From one to four species, the relationship is as presented in the main
text: total community \emph{CV} increases approximately linearly with
environmental variability because (1) environmental variability promotes
species coexistence \emph{and} (2) environmental variability causes
populations fluctuations to increase (Figure SI-4). However, after four
species, the relationship saturates --- species additions due to
coexistence by the storage effect completely buffer ecosystem
variability from further increases in environmental variability (Figure
SI-4). Thus, our results provide novel theoretical explanations for
positive and flat diversity--ecosystem variability relationships.
